---
title: "DATA 607 Assignment 1"
output: html_document
---

**Gehad Gad**

**February 2nd, 2020**

**Assignment 1**


**Introduction**
This data is about Breast Cancer findings, obtained from the University Medical Centre, Institute of Oncology, Ljubljana, Yugoslavia. 
This data set includes 201 instances of one class and 85 instances of another class. The instances are described by 9 attributes, some of which are linear and some are nominal. 


```{r}
#Import libraries and/or Packages
if (!require(dplyr)){
install.packages("dplyr")
library(dplyr)}
if (!require(caTools)){
install.packages("caTools")
library(caTools)}
if (!require(ggplot2)){
install.packages("ggplot2")
library(ggplot2)}

```


Read the data into R.Name the features of the data.

```{r}

BreastCancer <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer/breast-cancer.data", header=FALSE)

names (BreastCancer)<-c("output","age","menopause","tumor_size","inv_nodes","node_caps", "deg_malig","breast","breast_quad","irradiat")

```


Display few rows from the data.

```{r}
head(BreastCancer)
```

```{r}
#Display the head of the data.

summary(BreastCancer)
```


```{r}
#check for NA values in the data.

sum(is.na(BreastCancer))
```
Also, there is not NA showing, there are some ? in the date appears in the summary view.


```{r}
#Replace ? values in node_caps columns:

BreastCancer$node_caps[BreastCancer$node_caps=="?"]="yes"

```


```{r}
#Replace any NA's values in node_caps

BreastCancer$node_caps=ifelse(is.na(BreastCancer$node_caps),ave(BreastCancer$node_caps,FUN=function(x)"no"),BreastCancer$node_caps)
```


```{r}
#Replace any ? values in breast_quad column.

BreastCancer$breast_quad[BreastCancer$breast_quad=="?"]="left_low"
```


```{r}
#Replace NA's values in left_low columns:

BreastCancer$breast_quad=ifelse(is.na(BreastCancer$breast_quad),ave(BreastCancer$breast_quad,FUN=function(x)"left_low"),BreastCancer$breast_quad)
```

```{r}
#Check the levels of any any column

factor (BreastCancer$irradiat)
```


```{r}
#Labeling the features


BreastCancer$output= factor (BreastCancer$output, labels= c(0,1) , levels= c("no-recurrence-events", "recurrence-events"))

BreastCancer$age= factor (BreastCancer$age,labels= c(0,1,2,3,4,5) , levels= c("20-29","30-39","40-49","50-59","60-69","70-79")) 

BreastCancer$menopause= factor (BreastCancer$menopause, labels= c(0,1,2) , levels= c("premeno","ge40","lt40")) 

BreastCancer$tumor_size= factor (BreastCancer$tumor_size, labels= c(0,1,2,3,4,5,6,7,8,9,10) , levels= c("0-4", "10-14", "15-19", "20-24","25-29", "30-34", "35-39","40-44","45-49","5-9","50-54")) 

BreastCancer$inv_nodes= factor (BreastCancer$inv_nodes, labels= c(0,1,2,3,4,5,6) , levels= c("0-2", "12-14", "15-17", "24-26", "3-5", "6-8", "9-11"))


BreastCancer$node_caps= factor (BreastCancer$node_caps, labels= c(0,1) , levels= c("2", "3"))

BreastCancer$deg_malig= factor (BreastCancer$deg_malig, labels= c(0,1,2) , levels= c("1", "2", "3"))


BreastCancer$breast= factor (BreastCancer$breast, labels= c(0,1) , levels= c("left", "right")) 

BreastCancer$breast_quad= factor (BreastCancer$breast_quad, labels= c(0,1,2,3,4) , levels= c("2", "3", "4", "5", "6")) 

BreastCancer$irradiat= factor (BreastCancer$irradiat, labels= c(0,1) , levels= c("yes", "no")) 

```

```{r}

#Factorizing all features

FeaturesNames = list ("output","age","menopause","tumor_size","inv_nodes","node_caps", "deg_malig","breast","breast_quad","irradiat")

for(i in FeaturesNames){
  BreastCancer[i] <- sapply (BreastCancer[i], as.factor)
  
}
  
  
```


```{r}
#Split the data to input-output

y= select(BreastCancer,c(1))

x= select(BreastCancer,-c(output))

```


```{r}
#Split the data to train and test data for more analysis and modeling.

sample = sample.split(BreastCancer,SplitRatio = .75)
Train = subset (BreastCancer, sample ==TRUE) 
Test =  subset (BreastCancer, sample ==FALSE)

```



```{r}
ggplot(data=BreastCancer,aes(x=age,fill=irradiat,))+geom_bar()
```

The graph above shows that ages between 20-29 did not expose to irriadiation, ages between 70-79 expose just little. But the majority of ages between 30 to 69 shows more exposure to irradiation. 


```{r}
ggplot(BreastCancer,aes(x= age,fill= output)) +
  theme_bw() +
  facet_wrap(~menopause) +
  geom_bar()+
  labs(y ="Freq",
       title = "The Irradiat distribution by Age based on Menopause")

```

The graph above shows that distribution of irradiat based on age and menopause.



**Conclusion**

Data cleaning and processing is done. Since this data is classification, decision Tree can help us to visualize decision rules for predicting a categorical (classification tree).

